---
title: "Testicular cell atlas"
author: "Jarning"
date: "2022/1/1"
output: html_document
---

## setup code

```{r setup}
pacman::p_unload(pacman::p_loaded(), character.only = TRUE)
options(stringsAsFactors = FALSE)
library(here)
library(Seurat, lib.loc = "/home/software/R/seuratV3_env/")
library(SeuratWrappers)
library(tidyverse)
source("R/utils.R")
source("R/02.preprocess.utils.R")

# global vars
OUTDIR <- "../results/02.preprocess_mouse"
safe_mkdir(OUTDIR)

# check all mouse datasets
datasets <- list.dirs("../data/matrix", recursive = F, full.names = F)
datasets[grep("*mouse*", datasets)]
```


## Cell.Rep_2018_mouse_GSE109033

### 1) load data
```{r}
gse_dataset = "Cell.Rep_2018_mouse_GSE109033"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)


# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
## remove bad clusters
drop.clusters <- c(0,4,5,8,9,15,21) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)
DimPlot(seu, group.by = "RNA_snn_res.2", label = T)
drop.clusters <- c(43,40,39) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

seu$seurat_clusters <- NULL

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Cell.Rep_2018_mouse_GSE109033"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Cell.Res_2018_mouse_GSE107644

### 1) load data
```{r}
gse_dataset = "Cell.Res_2018_mouse_GSE107644"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 1000 & percent.mt <= 10)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```


### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

## add published cell type
pub.ct <- read.table(paste(path, "cell_type_published.txt", sep = "/"), 
                     sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))

metadata.new <- seu@meta.data

metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

cells.use <- rownames(subset(seu@meta.data, !is.na(Cell_type_published)))
seu <- subset(seu, cells = cells.use)
cells.use <- rownames(subset(seu@meta.data, Cell_type_published %notin% c("Sox30KORS2o3","Spo11KOL1")))
seu <- subset(seu, cells = cells.use)

# save data
obj.name <- paste0(seu@project.name, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Cell.Res_2018_mouse_GSE107644"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Dev.Cell_2018_mouse_GSE112393

### 1) load data
```{r}
gse_dataset = "Dev.Cell_2018_mouse_GSE112393"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

## add published cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))
metadata.new <- seu@meta.data
metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
## remove bad clusters
drop.clusters <- c(15) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(34,21,9,23,14,17,28) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 10, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(33) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

seu$seurat_clusters <- NULL
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Dev.Cell_2018_mouse_GSE112393"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Nat.Com_2019_mouse_E-MTAB-6946

### 1) load data
```{r}
gse_dataset = "Nat.Com_2019_mouse_E-MTAB-6946"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

## add published cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))

metadata.new <- seu@meta.data

metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

seu <- subset(seu, Genotype == "WT")

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
DimPlot(seu, group.by = "Cell_type_published", label = T) + NoLegend()
```

### 4) step1 clustering

```{r}
source("03.remove_doublets.R")

gse_dataset = "Nat.Com_2019_mouse_E-MTAB-6946"
obj.name <- paste0(gse_dataset, ".seu.rds")
seu <- readRDS(file.path(result.path, obj.name))

## remove bad clusters
drop.clusters <- c(6,19,21,25) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(50,56,53,46) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(54) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

seu$seurat_clusters <- NULL
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```


```{r}
gse_dataset = "Nat.Com_2019_mouse_E-MTAB-6946"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Nat.Com_2019_mouse_GSE124904

### 1) load data
```{r}
gse_dataset = "Nat.Com_2019_mouse_GSE124904"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
##
Idents(seu) <- seu$RNA_snn_res.1
markers.1 <- mcFindAllMarkers(seu, n.cores = 20)

## remove bad clusters
drop.clusters <- c(7,10) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

seu$seurat_clusters <- NULL
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Nat.Com_2019_mouse_GSE124904"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Plos.Gen_2019_mouse_GSE121904

### 1) load data
```{r}
gse_dataset = "Plos.Gen_2019_mouse_GSE121904"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
##
Idents(seu) <- seu$RNA_snn_res.1
markers.1 <- mcFindAllMarkers(seu, n.cores = 20)

## remove bad clusters
drop.clusters <- c(1,14,20) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(22,25,42,47,46) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30, n.neighbors = 100)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(10,16) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30, n.neighbors = 100)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(41) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

seu$seurat_clusters <- NULL
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Plos.Gen_2019_mouse_GSE121904"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Elife_2019_mouse_GSE113293

### 1) load data
```{r}
gse_dataset = "Elife_2019_mouse_GSE113293"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)
seu <- subset(seu, Genotype == "WT")

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
##
Idents(seu) <- seu$RNA_snn_res.1
markers.1 <- FindMarkers(seu, ident.1 = 1, only.pos = T)

## remove bad clusters
drop.clusters <- c(1) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
# seu <- DF_main(seu, split.by = "GSM_ID")
# seu <- subset(seu, DF.classifications == "Singlet")
# seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

seu$seurat_clusters <- NULL
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Elife_2019_mouse_GSE113293"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Dev_2020_mouse_GSE130593

### 1) load data
```{r}
gse_dataset = "Dev_2020_mouse_GSE130593"
seu <- load_dataset(gse_dataset, species = "mouse")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
##
Idents(seu) <- seu$RNA_snn_res.1
markers.1 <- FindMarkers(seu, ident.1 = 1, only.pos = T)

## remove bad clusters
drop.clusters <- c() # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.1 %notin% drop.clusters)
seu <- subset(seu, percent.mt <= 10)

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

drop.clusters <- c(25,32) # high percent.mt and low genes, multiple markers
seu <- subset(seu, RNA_snn_res.2 %notin% drop.clusters)

## re-run the seurat pipeline
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
seu <- seu %>% 
    FindNeighbors(dims = 1:30, reduction = "mnn", k.param = 20, verbose = FALSE) %>% 
    FindClusters(resolution = 2, verbose = FALSE)

seu$seurat_clusters <- NULL
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Dev_2020_mouse_GSE130593"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Nat.Com_2021_mouse_GSE148032

### 1) load data
```{r}
gse_dataset = "Nat.Com_2021_mouse_GSE148032"
seu <- load_dataset(gse_dataset, species = "mouse")
seu$orig.ident <- sub("_[ATGC]+$", "", colnames(seu))
Idents(seu) <- seu$orig.ident
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 1000 & percent.mt <= 10)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = result.path)

# add metadata
seu <- add_metadata(seu)

# add public cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))
metadata.new <- seu@meta.data
metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

# add public cell meta
pub.meta <- read.table(paste(INDIR, "cellmeta.txt", sep = "/"), 
                       sep = "\t", header = F,
                       col.names = c("CellID", "Age", "Sex", "Genotype", "Cell_enrichment"))
metadata.new$Age <- mapvalues.2(x = rownames(metadata.new), from = pub.meta$CellID, to = pub.meta$Age)
metadata.new$Sex <- mapvalues.2(x = rownames(metadata.new), from = pub.meta$CellID, to = pub.meta$Sex)
metadata.new$Genotype <- mapvalues.2(x = rownames(metadata.new), from = pub.meta$CellID, to = pub.meta$Genotype)
metadata.new$Cell_enrichment <- mapvalues.2(x = rownames(metadata.new), from = pub.meta$CellID, to = pub.meta$Cell_enrichment)
seu@meta.data <- metadata.new

seu <- subset(seu, Genotype == "WT")

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
seu <- subset(seu, cells = rownames(subset(seu@meta.data, !is.na(Cell_type_published))) )
seu <- subset(seu, Cell_type_published != "Outliers")

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

```{r}
gse_dataset = "Nat.Com_2021_mouse_GSE148032"
seu.raw <- load_dataset(gse_dataset, species = "mouse")

## add new cell meta
obj.name <- paste0(gse_dataset, ".seu.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
seu <- readRDS(file.path(OUTDIR, obj.name))
new.meta <- seu@meta.data
new.meta <- new.meta |>
  mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)],
         percent.chrY = seu.raw$percent.chrY[rownames(new.meta)],
         decontX_contamination = seu.raw$decontX_contamination[rownames(new.meta)],
         .after="percent.ribo")
seu@meta.data <- new.meta
saveRDS(seu, file.path(OUTDIR, obj.name))
rm("seu.raw")
```

## Merge datasets

```{r}
datasets <- list.dirs("../data/matrix", recursive = F, full.names = F)
datasets <- datasets[grep("*mouse*", datasets)]

seu.merged <- lapply(datasets, function(xx) load_processed_dataset(xx, species = "mouse")) %>% reduce(., merge)

sceasy::convertFormat(seu.merged, from="seurat", to="anndata", main_layer = "counts",
                      outFile=file.path(OUTDIR, 'TCA_mouse.processed.h5ad'))
```
