---
title: "Testicular cell atlas"
author: "Jarning"
date: "2021/12/30"
output: html_document
---

## setup code

```{r setup}
pacman::p_unload(pacman::p_loaded(), character.only = TRUE)
options(stringsAsFactors = FALSE)
library(here)
library(Seurat, lib.loc = "/home/software/R/seuratV3_env/")
library(SeuratWrappers)
library(tidyverse)
source("R/utils.R")
source("R/02.preprocess.utils.R")

# global vars
OUTDIR <- "../results/02.preprocess_human"
safe_mkdir(OUTDIR)

# check all human datasets
datasets <- list.dirs("../data/matrix", recursive = F, full.names = F)
datasets[grep("*human*", datasets)]
```

## Cell.Stem.Cell_2017_human_GSE86146

### 1) load data
```{r}
gse_dataset = "Cell.Stem.Cell_2017_human_GSE86146"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 1000 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

## add published cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type","age","sex"))
metadata.new <- seu@meta.data
metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
metadata.new$Age <- mapvalues.2(x = rownames(metadata.new), 
                                from = pub.ct$cellID, 
                                to = pub.ct$age)
metadata.new$Sex <- mapvalues.2(x = rownames(metadata.new), 
                                from = pub.ct$cellID, 
                                to = pub.ct$sex)
seu@meta.data <- metadata.new

# filter cells
seu <- subset(seu, Sex == "Male")

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Cell_type_published", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Cell.Stem.Cell_2017_human_GSE86146"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)
obj.name <- paste0(gse_dataset, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```


## Cell.Rep_2018_human_GSE109037

### 1) load data
```{r}
gse_dataset = "Cell.Rep_2018_human_GSE109037"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR, do.plot = TRUE)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Cell_enrichment", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Cell.Rep_2018_human_GSE109037"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# obj.name
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## filter cells
DimPlot(seu, group.by = "RNA_snn_res.1", label = T)
FeaturePlot(seu, features = "decontX_contamination")
seu <- subset(seu, RNA_snn_res.1 %notin% c(1,9,14))

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
saveRDS(seu, file.path(OUTDIR, obj.name))
```

## Cell.Res_2018_human_GSE120508

### 1) load data
```{r}
gse_dataset = "Cell.Res_2018_human_GSE120508"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

# add public cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))

metadata.new <- seu@meta.data
metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Cell_type_published", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Cell.Stem.Cell_2018_human_GSE106487"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# obj.name
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## filter cells
DimPlot(seu, group.by = "RNA_snn_res.1", label = T)
FeaturePlot(seu, features = "decontX_contamination")
seu <- subset(seu, Cell_published %notin% c(3,11))

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
saveRDS(seu, file.path(OUTDIR, obj.name))
```

## Cell.Stem.Cell_2018_human_GSE106487

### 1) load data
```{r}
gse_dataset = "Cell.Stem.Cell_2018_human_GSE106487"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 1000 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

# add public cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))

metadata.new <- seu@meta.data
metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Cell_type_published", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.hc.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Cell.Res_2018_human_GSE120508"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# obj.name
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## filter cells
DimPlot(seu, group.by = "RNA_snn_res.1", label = T)
FeaturePlot(seu, features = "decontX_contamination")
kept.cells <- rownames(subset(seu@meta.data, !is.na(seu$Cell_type_published)))
seu <- subset(seu, cells = kept.cells)

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
saveRDS(seu, file.path(OUTDIR, obj.name))
```

## Cell.Rep_2019_human_GSE124263

### 1) load data
```{r}
gse_dataset = "Cell.Rep_2019_human_GSE124263"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Cell.Rep_2019_human_GSE124263"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# obj.name
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## filter cells
DimPlot(seu, group.by = "RNA_snn_res.1", label = T)
FeaturePlot(seu, features = "decontX_contamination")
## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
DimPlot(seu, group.by = "DF.classifications", label = T)
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
saveRDS(seu, file.path(OUTDIR, obj.name))
```

## Cell.Stem.Cell_2020_human_GSE134144

### 1) load data
```{r}
gse_dataset = "Cell.Stem.Cell_2020_human_GSE134144"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

# add public cell type
INDIR = here("data/matrix", gse_dataset)
pub.ct <- read.table(file.path(INDIR, "cell_type_published.txt"), sep = "\t", header = F, 
                     col.names = c("cellID", "cell_type"))
metadata.new <- seu@meta.data

metadata.new$Cell_type_published <- mapvalues.2(x = rownames(metadata.new), 
                                                from = pub.ct$cellID, 
                                                to = pub.ct$cell_type)
seu@meta.data <- metadata.new

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Cell.Stem.Cell_2020_human_GSE134144"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# obj.name
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## filter cells
DimPlot(seu, group.by = "RNA_snn_res.1", label = T)
FeaturePlot(seu, features = "decontX_contamination")

## remove doublets
seu <- DF_main(seu, split.by = "GSM_ID")
DimPlot(seu, group.by = "DF.classifications", label = T)
seu <- subset(seu, DF.classifications == "Singlet")
seu$DF.classifications <- NULL

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
saveRDS(seu, file.path(OUTDIR, obj.name))
```

## Nat.Com_2020_human_GSE149512

### 1) load data
```{r}
gse_dataset = "Nat.Com_2020_human_GSE149512"
seu <- load_dataset(gse_dataset, species = "human")
```

### 2) filter cells
```{r, fig.width=20, fig.height=9}
VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = T)

seu <- subset(seu, nFeature_RNA >= 500 & percent.mt <= 15)

VlnPlot(object = seu, 
        features = c("nFeature_RNA", "percent.mt", "percent.ribo"), 
        ncol = 2, pt.size = 0, log = F)
```

### 3) add cell metadata
```{r}
# run umap and cluster
seu <- seurat_one_step(seu, outpath = OUTDIR)

# add metadata
seu <- add_metadata(seu)

# show embeddings
FeaturePlot(seu, features = "decontX_contamination")
DimPlot(seu, group.by = "Sample_name", label = T)
DimPlot(seu, label = T)

# save data
obj.name <- paste0(seu@project.name, ".seu.rds")
saveRDS(seu, file.path(OUTDIR, obj.name))
```

### 4) step1 clustering

```{r}
# gse_dataset = "Nat.Com_2020_human_GSE149512"
# seu.raw <- load_dataset(gse_dataset, species = "human")
# 
# ## add new cell meta
# obj.name <- paste0(gse_dataset, ".seu.rds")
# obj.name
# seu <- readRDS(file.path(OUTDIR, obj.name))
# new.meta <- seu@meta.data
# new.meta <- new.meta |> 
#   mutate(percent.chrX = seu.raw$percent.chrX[rownames(new.meta)], 
#          percent.chrY = seu.raw$percent.chrY[rownames(new.meta)], 
#          .before="decontX_contamination")
# seu@meta.data <- new.meta
# 
# saveRDS(seu, file.path(OUTDIR, obj.name))
# rm("seu.raw")

## filter cells
DimPlot(seu, group.by = "RNA_snn_res.1", label = T)
FeaturePlot(seu, features = "decontX_contamination")

## remove doublets
seu <- subset(seu, RNA_snn_res.1 %notin% c(3))

## redo umap
seu <- RunUMAP(seu, reduction = "mnn", dims = 1:30)

obj.name <- paste0(gse_dataset, ".seu.hc.rds")
obj.name
saveRDS(seu, file.path(OUTDIR, obj.name))
```

## Merge datasets

```{r}
seu.merged <- list(
  "Li.2017" = load_processed_dataset("Cell.Stem.Cell_2017_human_GSE86146"),
  "Hermann.2018" = load_processed_dataset("Cell.Rep_2018_human_GSE109037"),
  "Guo.2018" = load_processed_dataset("Cell.Res_2018_human_GSE120508"),
  "Wang.2018" = load_processed_dataset("Cell.Stem.Cell_2018_human_GSE106487"),
  "Sohni.2019" = load_processed_dataset("Cell.Rep_2019_human_GSE124263"),
  "Guo.2020" = load_processed_dataset("Cell.Stem.Cell_2020_human_GSE134144"),
  "Zhao.2020" = load_processed_dataset("Nat.Com_2020_human_GSE149512")
) %>% reduce(., merge)

sceasy::convertFormat(seu.merged, from="seurat", to="anndata", main_layer = "counts",
                      outFile=file.path(OUTDIR, 'TCA_human.processed.h5ad'))
```

